---
title: "Bellabeat Case Study Report"
author: "Albert Hunduza"
date: "2023-08-28"
output: html_document
---



## R Markdown

In this notebook, I outline how I approached the Bellabeat case study as part of my Google Data Analytics Capstone Project. The full description of the case study is attached in the file xxxxxxx. A brief outline of the case study scenario is that I am working on the marketing analyst team at Bellabeat,  a high-tech manufacturer of
health-focused products for women. The goal is to analyze smart device usage data in order to gain insight into how people are already using their smart devices. Using these insights, I have to give high-level recommendations for how these trends can inform Bellabeat marketing strategy, particularly for one of the Bellabeat products. Key questions include, but are not limited to: <br>

1. What are some trends in smart device usage? <br>
2. How could these trends apply to Bellabeat customers? <br>
3. How could these trends help influence Bellabeat marketing strategy? <br>

This study is one of the key deliverables, mostly relevant to my colleageues on the marketing data analytics team at Bellabeat. I started by importing the required libraries.

```{r}
# Load the required libraries
library(dplyr)
library(ggplot2)
library(corrplot)
library(gridExtra)
```

The next step was to import the relevant datasets from the dataset collection. I limited my choices to datasets that recorded daily and hourly data as these are more informative to a higher level marketing strategy.

```{r}
# Importing relevant datasets
data_folder_path <- "Fitabase Data 4.12.16-5.12.16/" # sets path of folder containing files

daily_activity <- read.csv(paste(data_folder_path,"dailyActivity_merged.csv",sep=""))
sleep_day <- read.csv(paste(data_folder_path,"sleepDay_merged.csv",sep=""))
weight_log <- read.csv(paste(data_folder_path,"weightLogInfo_merged.csv",sep=""))
hourly_intensities <- read.csv(paste(data_folder_path,"hourlyIntensities_merged.csv",sep=""))
hourly_steps <- read.csv(paste(data_folder_path,"hourlySteps_merged.csv",sep=""))
heart_rate <- read.csv(paste(data_folder_path,"heartrate_seconds_merged.csv",sep=""))

```

I started by cleaning the data by checking for NaNs and duplicates:

```{r}
# Checking for NaNs in each dataset
n_nans_daily_activity <- colSums(is.na(daily_activity))
n_nans_sleep_day <- colSums(is.na(sleep_day))
n_nans_weight_log <- colSums(is.na(weight_log))
n_nans_hourly_intensities <- colSums(is.na(hourly_intensities))
n_nans_hourly_steps <- colSums(is.na(hourly_steps))
n_nans_heart_rate <- colSums(is.na(heart_rate)) 

# Print the count of NaNs for each dataset
n_nans_daily_activity
n_nans_sleep_day
n_nans_weight_log
n_nans_hourly_intensities
n_nans_hourly_steps
n_nans_heart_rate

```

Only the weight_log dataframe contained null values, and these were in the "Fat" column where some people in the study did not report their measurements for fat. Since an overwhelming majority (65/67) of records had null values in this column, I could not drop it. I therefore replaced all the NaNs in this column with "unreported"

```{r}
# Replace NaNs in the "Fat" column with "unreported"
weight_log$Fat <- ifelse(is.na(weight_log$Fat), "unreported", weight_log$Fat)

# checking if NaNs have been dropped
n_nans_weight_log <- colSums(is.na(weight_log))
n_nans_weight_log

```

Having eliminated the NaNs, I checked for duplicates

```{r}
# Checking for duplicate rows in each dataset
n_duplicates_daily_activity <- sum(duplicated(daily_activity))
n_duplicates_sleep_day <- sum(duplicated(sleep_day))
n_duplicates_weight_log <- sum(duplicated(weight_log))
n_duplicates_hourly_intensities <- sum(duplicated(hourly_intensities))
n_duplicates_hourly_steps <- sum(duplicated(hourly_steps))
n_duplicates_heart_rate <- sum(duplicated(heart_rate))

# Print the count of duplicate rows for each dataset
n_duplicates_daily_activity
n_duplicates_sleep_day
n_duplicates_weight_log
n_duplicates_hourly_intensities
n_duplicates_hourly_steps
n_duplicates_heart_rate

```

The sleep_day dataframe had 3 duplicates. I eliminated them as follows:

```{r}
# Remove duplicate rows from the sleep_day dataframe
sleep_day_cleaned <- sleep_day[!duplicated(sleep_day), ]

# check if duplicates have been removed
n_duplicates_sleep_day_cleaned <- sum(duplicated(sleep_day_cleaned))
n_duplicates_sleep_day_cleaned

```

The hourly_intensities and hourly_steps dataframes are narrow and contain information that is related to each other. I therefore joined them on the Id and date columns into a new dataframe.

```{r}
# merging the hourly_intensities and hourly_steps dataframes
hourly_exercises <- merge(hourly_intensities, hourly_steps, by = c("Id", "ActivityHour"))

```

My approach was to analyze each dataframe separately and etablish trends and behaviors from each before finally looking for insights across the dataframes

### 1) daily_activity dataframe

```{r}
# Get the number of unique IDs/people in the "Id" column
unique_id_count <- daily_activity %>%
  distinct(Id) %>%
  nrow()

# Print the number of unique IDs/people
print(unique_id_count)

```


There are 33 people who recorded their daily activity. Next we will find how many days each person recorded their activity. This will help us establish how useful our data can be.

```{r}
# Convert the "ActivityDate" column to Date format
daily_activity %>%
  mutate(ActivityDate = as.Date(ActivityDate, format = "%m/%d/%Y")) %>%
  distinct(ActivityDate) %>%
  summarise(unique_date_count = n())



# Convert the "ActivityDate" column to Date format and group by Id
result <- daily_activity %>%
  mutate(ActivityDate = as.Date(ActivityDate, format = "%m/%d/%Y")) %>%
  group_by(Id) %>%
  summarise(unique_date_count = n_distinct(ActivityDate))

# Set the option to display all rows
options(dplyr.print_max = Inf)

# Print the result
print(result)

```

Not all participants recorded their daily activities on all 31 days of the study. One of the participants only recorded their data on 4 days.
Next, I calculated the average number of steps and distance that each person walked over this study period.

```{r}
# Calculate the average number of steps and distance per person
average_per_person <- daily_activity %>%
  group_by(Id) %>%
  summarise(average_steps = mean(TotalSteps, na.rm = TRUE),
            average_distance = mean(TotalDistance, na.rm = TRUE))

# Print the result
print(average_per_person)

```

These results are best visualized in a histogram

```{r}
# Calculate the average number of steps and distance per person
average_per_person <- daily_activity %>%
  group_by(Id) %>%
  summarise(average_steps = mean(TotalSteps, na.rm = TRUE),
            average_distance = mean(TotalDistance, na.rm = TRUE))

# Create a histogram for average steps
ggplot(average_per_person, aes(x = average_steps)) +
  geom_histogram(binwidth = 500) +
  labs(title = "Histogram of Average Steps per Person",
       x = "Average Steps",
       y = "Frequency")

# Create a histogram for average distance
ggplot(average_per_person, aes(x = average_distance)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Histogram of Average Distance per Person",
       x = "Average Distance",
       y = "Frequency")

```

On average, people in the study walked about 8000 steps or around 6km a day, with 16000 steps being the high end and 1000 the lower.

Next I created a correlation matrix to see which factor influences the amount of calories burned more between total steps and total distance.

```{r}
# Select the relevant columns
selected_columns <- daily_activity %>%
  select(TotalSteps, TotalDistance, Calories)

# Calculate the correlation matrix
correlation_matrix <- cor(selected_columns, use = "complete.obs")

# Visualize the correlation matrix with a title
corrplot(correlation_matrix, method = "shade", type = "upper",
         tl.col = "black", tl.srt = 45)

```

From the correlation matrix, we can see that total distance moved has a slightly greater influence on calories burned than total steps. Both factors are positively correlated to calories burned.

Next, I sought to discover if there are any days of the week when people tend to be more active than others


```{r}
# Convert the "ActivityDate" column to Date format and add a new column for the day of the week
daily_activity_with_day <- daily_activity %>%
  mutate(ActivityDate = as.Date(ActivityDate, format = "%m/%d/%Y"),
         DayOfWeek = weekdays(ActivityDate))

# Group by day of the week and calculate the average distance
average_distance_per_day <- daily_activity_with_day %>%
  group_by(DayOfWeek) %>%
  summarise(average_distance = mean(TotalDistance, na.rm = TRUE))

# Order the days of the week
day_order <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
average_distance_per_day$DayOfWeek <- factor(average_distance_per_day$DayOfWeek, levels = day_order)

# Create a bar plot
ggplot(average_distance_per_day, aes(x = DayOfWeek, y = average_distance, fill = DayOfWeek)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Distance per Day of the Week",
       x = "Day of the Week",
       y = "Average Distance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The average distance moved seems to be constant for all days of the week. Saturday and Tuesday have slightly more distance than the other days and Sunday slightly less.

Another important thing was to see how the different intensities of activity influence calories burned. I used another correlation matrix to visualize this.

```{r}
# Select the relevant columns
selected_columns <- daily_activity %>%
  select(VeryActiveDistance, ModeratelyActiveDistance, LightActiveDistance, Calories)

# Calculate the correlation matrix
correlation_matrix <- cor(selected_columns, use = "complete.obs")

# Visualize the correlation matrix with a title
corrplot(correlation_matrix, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45)

```

I did the same thing for minutes

```{r}
# Select the relevant columns
selected_columns <- daily_activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes, Calories)

# Calculate the correlation matrix
correlation_matrix <- cor(selected_columns, use = "complete.obs")

# Visualize the correlation matrix with a title
corrplot(correlation_matrix, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45)

```

From these correlation matrices, we can see that Very Intense activity has the greatest effect on calories burned. Moderate and light activity are also positively correlated to calories burned while sedentary minutes are negatively correlated to calories burned.

### 2) sleep_day dataframe

I started by counting the number of people in this dataset and how many contibutions they each had

```{r}
# Number of unique people (Ids)
unique_people_count <- sleep_day_cleaned %>%
  summarise(unique_people = n_distinct(Id))

# Print the number of unique people
print(unique_people_count)

# Number of unique days each person has a record for
unique_days_per_person <- sleep_day_cleaned %>%
  group_by(Id) %>%
  summarise(unique_days = n_distinct(as.Date(SleepDay, format = "%m/%d/%Y")))

# Print the number of unique days per person
print(unique_days_per_person)

```


Next, I wanted to find out how many times people tend to sleep per day

```{r}
# Calculate the percentages
percentage_data <- sleep_day_cleaned %>%
  group_by(TotalSleepRecords) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)

# Create a bar graph with percentages
ggplot(percentage_data, aes(x = factor(TotalSleepRecords), y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), vjust = -0.5, size = 3, color = "black") +
  labs(title = "Distribution of TotalSleepRecords",
       x = "Total Sleep Records",
       y = "Percentage") +
  theme_minimal() +
  scale_x_discrete(labels = function(x) sprintf("%d", as.integer(x))) 

```

89% of the time, people in the study slept just once per day, 10% of the times they slept twice and very rarely slept thrice

```{r}
# Calculate the average TotalSleepRecords per unique Id
average_sleep_records_per_id <- sleep_day_cleaned %>%
  group_by(Id) %>%
  summarise(average_sleep_records = mean(TotalSleepRecords, na.rm = TRUE))

# Print the result
print(average_sleep_records_per_id)

# Create a boxplot for TotalMinutesAsleep
ggplot(sleep_day_cleaned, aes(y = TotalMinutesAsleep)) +
  geom_boxplot(fill = "blue", color = "black") +
  labs(title = "Boxplot of TotalMinutesAsleep",
       y = "Total Minutes Asleep") +
  theme_minimal()

```

This is corroborated by these averages that shows that each person slept only once on average, with 1.23 being the maximum average number of sleep records per person.

```{r}
# Calculate the average TotalMinutesAsleep
average_minutes_asleep <- mean(sleep_day_cleaned$TotalMinutesAsleep, na.rm = TRUE)

# Print the result
print(average_minutes_asleep)


# Calculate the average TotalMinutesAsleep per person
average_minutes_asleep_per_person <- sleep_day_cleaned %>%
  group_by(Id) %>%
  summarise(average_minutes_asleep = mean(TotalMinutesAsleep, na.rm = TRUE))

# Print the result
print(average_minutes_asleep_per_person)

```
The average sleep time is 419 minutes which is about 7 hours.

Next, I calculated the amount of time spent in bed actually sleeping.

```{r}
# Calculate the percentage of time spent in bed that each person was actually asleep
percentage_time_asleep_per_person <- sleep_day_cleaned %>%
  mutate(PercentageAsleep = (TotalMinutesAsleep / TotalTimeInBed) * 100) %>%
  group_by(Id) %>%
  summarise(AveragePercentageAsleep = mean(PercentageAsleep, na.rm = TRUE))

# Print the result
print(percentage_time_asleep_per_person)

```
This is important because it hints at sleep quality. Most of the participants were asleep more than 90% of the time they were in bed on average. The highest was 98.5% and the lowest was 67.8%


```{r}
# Convert the "SleepDay" column to Date format and add a new column for the day of the week
sleep_day_cleaned <- sleep_day_cleaned %>%
  mutate(SleepDay = as.Date(SleepDay, format = "%m/%d/%Y"),
         DayOfWeek = weekdays(SleepDay))

# Calculate the average TotalMinutesAsleep and percentage of Time in Bed spent asleep for each day of the week
average_sleep_per_day_of_week <- sleep_day_cleaned %>%
  group_by(DayOfWeek) %>%
  summarise(AverageTotalMinutesAsleep = mean(TotalMinutesAsleep, na.rm = TRUE),
            AveragePercentageAsleep = mean((TotalMinutesAsleep / TotalTimeInBed) * 100, na.rm = TRUE))

# Create bar graphs
ggplot(average_sleep_per_day_of_week, aes(x = DayOfWeek)) +
  geom_bar(aes(y = AverageTotalMinutesAsleep), stat = "identity", fill = "blue", alpha = 0.6, position = "dodge") +
  geom_bar(aes(y = AveragePercentageAsleep), stat = "identity", fill = "green", alpha = 0.6, position = "dodge") +
  labs(title = "Average Total Minutes Asleep and % of Time in Bed Spent Asleep by DOW",
       x = "Day of the Week",
       y = "Value") +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = "Percentage of Time in Bed Spent Asleep")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("blue", "green")) +
  scale_alpha_manual(values = c(0.6, 0.6)) +
  guides(fill = guide_legend(title = "Legend", override.aes = list(alpha = 1)))

```

Again, there doesn't seem to be any significant differences between hours spent asleep on different days of the week.On average the people in the study slept slightly more on Sundays than on the other days of the week.

### 3) weight_log dataframe

```{r}
# Number of unique people (Ids)
unique_people_count <- weight_log %>%
  summarise(unique_people = n_distinct(Id))

# Print the number of unique people
print(unique_people_count)

# Total days with records for each person
total_days_per_person <- weight_log %>%
  group_by(Id) %>%
  summarise(total_days = n_distinct(Date))

# Print the total days with records per person
print(total_days_per_person)

```

We can see from the results that much fewer people recorded their weight during the study and even fewer did it consistently.

```{r}
# Convert the "Date" column to a proper date format
weight_log$Date <- as.POSIXct(weight_log$Date, format = "%m/%d/%Y %I:%M:%S %p")

# Find the starting weight for each person
starting_weight <- weight_log %>%
  group_by(Id) %>%
  summarise(starting_weight = first(WeightKg))

# Find the ending weight for each person
ending_weight <- weight_log %>%
  group_by(Id) %>%
  summarise(ending_weight = last(WeightKg))

# Combine the starting and ending weights
weight_summary <- left_join(starting_weight, ending_weight, by = "Id")

# Create a bar graph
ggplot(weight_summary, aes(x = factor(Id))) +
  geom_bar(aes(y = starting_weight), stat = "identity", fill = "blue", alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_bar(aes(y = ending_weight), stat = "identity", fill = "green", alpha = 0.6, position = position_dodge(width = 0.8)) +
  labs(title = "Starting and Ending Weights for Each Person",
       x = "ID",
       y = "Weight (Kg)") +
  scale_fill_manual(values = c("blue", "green")) +
  scale_alpha_manual(values = c(0.6, 0.6)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Over the month of the study, some of the participants lost a small amount of weight.However, the sample size is too small to try to infer a trend from.

```{r}
# Convert the "Date" column to a proper date format
weight_log$Date <- as.POSIXct(weight_log$Date, format = "%m/%d/%Y %I:%M:%S %p")

# Create a function to categorize BMI
categorize_bmi <- function(bmi) {
  if (bmi < 18.5) {
    return("Underweight")
  } else if (bmi >= 18.5 && bmi <= 25) {
    return("Normal")
  } else {
    return("Overweight")
  }
}

# Categorize BMI for the first and last records
weight_log_summary <- weight_log %>%
  group_by(Id) %>%
  arrange(Date) %>%
  summarise(StartCategory = categorize_bmi(first(BMI)),
            EndCategory = categorize_bmi(last(BMI)))

# Create pie charts for the start and end categories
pie_chart_start <- weight_log_summary %>%
  count(StartCategory) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ggplot(aes(x = "", y = n, fill = StartCategory, label = paste0(percentage, "%"))) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = paste0(percentage, "%")), position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("red", "green", "blue")) +  # Red for "Overweight," green for "Normal"
  labs(title = "Distribution of BMI Categories at Start") +
  theme_void() +
  theme(legend.position = "right")

pie_chart_end <- weight_log_summary %>%
  count(EndCategory) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ggplot(aes(x = "", y = n, fill = EndCategory, label = paste0(percentage, "%"))) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = paste0(percentage, "%")), position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("red", "green", "blue")) +  # Red for "Overweight," green for "Normal"
  labs(title = "Distribution of BMI Categories at End") +
  theme_void() +
  theme(legend.position = "right")

# Display the pie charts
gridExtra::grid.arrange(pie_chart_start, pie_chart_end, ncol = 2)

```

We can see that 37.5% of the participants were overweight and the rest were normal weight.

### 4) hourly_exercises dataframe

```{r}
# Number of unique people (Ids)
unique_people_count <- hourly_exercises %>%
  summarise(unique_people = n_distinct(Id))

# Print the number of unique people
print(unique_people_count)

# Number of days and unique hours per person
days_and_hours_per_person <- hourly_exercises %>%
  group_by(Id) %>%
  summarise(total_days = n_distinct(as.Date(ActivityHour)),
            unique_hours = n_distinct(ActivityHour))

# Print the number of days and unique hours per person
print(days_and_hours_per_person)

```


```{r}
# Calculate the average TotalIntensity and average StepTotal per person
average_intensity_and_steps_per_person <- hourly_exercises %>%
  group_by(Id) %>%
  summarise(average_total_intensity = mean(TotalIntensity, na.rm = TRUE),
            average_step_total = mean(StepTotal, na.rm = TRUE))

# Print the result
print(average_intensity_and_steps_per_person)

```

```{r}
# Calculate the average TotalIntensity and average StepTotal per person
average_intensity_and_steps_per_person <- hourly_exercises %>%
  group_by(Id) %>%
  summarise(average_total_intensity = mean(TotalIntensity, na.rm = TRUE),
            average_step_total = mean(StepTotal, na.rm = TRUE))

# Create a histogram for average TotalIntensity
histogram_intensity <- ggplot(average_intensity_and_steps_per_person, aes(x = average_total_intensity)) +
  geom_histogram(binwidth = 10, fill = "blue", alpha = 0.6) +
  labs(title = "Histogram of Average TotalIntensity per Person",
       x = "Average TotalIntensity",
       y = "Frequency") +
  theme_minimal()

# Create a histogram for average StepTotal
histogram_steps <- ggplot(average_intensity_and_steps_per_person, aes(x = average_step_total)) +
  geom_histogram(binwidth = 1000, fill = "green", alpha = 0.6) +
  labs(title = "Histogram of Average StepTotal per Person",
       x = "Average StepTotal",
       y = "Frequency") +
  theme_minimal()

# Display the histograms
gridExtra::grid.arrange(histogram_intensity, histogram_steps, ncol = 2)

```

From the histograms, we can see that the majority of people took 500 steps or less at high intensity during exercise.


```{r}
# Convert the "ActivityHour" column to a proper date format and extract the day of the week
hourly_exercises <- hourly_exercises %>%
  mutate(ActivityHour = as.POSIXct(ActivityHour, format = "%m/%d/%Y %I:%M:%S %p"),
         DayOfWeek = weekdays(ActivityHour))

# Calculate the average TotalIntensity and average StepTotal per day of the week
average_intensity_per_day <- hourly_exercises %>%
  group_by(DayOfWeek) %>%
  summarise(average_total_intensity = mean(TotalIntensity, na.rm = TRUE))

average_steps_per_day <- hourly_exercises %>%
  group_by(DayOfWeek) %>%
  summarise(average_step_total = mean(StepTotal, na.rm = TRUE))

# Create a bar plot for average TotalIntensity per day of the week
bar_plot_intensity <- ggplot(average_intensity_per_day, aes(x = DayOfWeek, y = average_total_intensity)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.6) +
  labs(title = "Average TotalIntensity per Day of the Week",
       x = "Day of the Week",
       y = "Average TotalIntensity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create a bar plot for average StepTotal per day of the week
bar_plot_steps <- ggplot(average_steps_per_day, aes(x = DayOfWeek, y = average_step_total)) +
  geom_bar(stat = "identity", fill = "green", alpha = 0.6) +
  labs(title = "Average StepTotal per Day of the Week",
       x = "Day of the Week",
       y = "Average StepTotal") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the bar plots
gridExtra::grid.arrange(bar_plot_intensity, bar_plot_steps, ncol = 2)

```

Again, there seems to be constant exercise over the week, except for Sunday which has slightly less than the rest and Saturday which has slightly more.

### 5) heart_rate dataframe

```{r}
# Number of unique people (Ids)
unique_people_count <- heart_rate %>%
  summarise(unique_people = n_distinct(Id))

# Print the number of unique people
print(unique_people_count)

# Number of unique records per person
unique_records_per_person <- heart_rate %>%
  group_by(Id) %>%
  summarise(unique_records = n_distinct(Time))

# Print the number of unique records per person
print(unique_records_per_person)

```

```{r}
# Calculate the average heart rate per person
average_heart_rate_per_person <- heart_rate %>%
  group_by(Id) %>%
  summarise(average_heart_rate = mean(Value, na.rm = TRUE))

# Create a bar plot for average heart rate per person
bar_plot_heart_rate <- ggplot(average_heart_rate_per_person, aes(x = factor(Id), y = average_heart_rate)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.6) +
  labs(title = "Average Heart Rate per Person",
       x = "ID",
       y = "Average Heart Rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the bar plot
print(bar_plot_heart_rate)

```


```{r}
# Convert the "Time" column to a proper datetime format
heart_rate$Time <- as.POSIXct(heart_rate$Time, format = "%m/%d/%Y %I:%M:%S %p")

# Extract the hour of the day from the "Time" column
heart_rate$HourOfDay <- format(heart_rate$Time, format = "%H")

# Calculate the average heart rate for each hour of the day
average_heart_rate_per_hour <- heart_rate %>%
  group_by(HourOfDay) %>%
  summarise(average_heart_rate = mean(Value, na.rm = TRUE))

# Create a line plot for average heart rate per hour of the day
line_plot_heart_rate <- ggplot(average_heart_rate_per_hour, aes(x = HourOfDay, y = average_heart_rate, group = 1)) +
  geom_line(color = "blue") +
  labs(title = "Average Heart Rate per Hour of the Day",
       x = "Hour of the Day",
       y = "Average Heart Rate") +
  theme_minimal()

# Display the line plot
print(line_plot_heart_rate)

```

### 6) Cross Dataframe analysis

```{r}
# Create a data frame with the number of distinct Ids for each dataframe
distinct_ids <- data.frame(
  DataFrame = c("daily_activity", "sleep_day_cleaned", "weight_log", "hourly_exercises", "heart_rate"),
  DistinctIds = c(
    n_distinct(daily_activity$Id),
    n_distinct(sleep_day_cleaned$Id),
    n_distinct(weight_log$Id),
    n_distinct(hourly_exercises$Id),
    n_distinct(heart_rate$Id)
  )
)

# Create a bar graph
bar_graph <- ggplot(distinct_ids, aes(x = DataFrame, y = DistinctIds, fill = DataFrame)) +
  geom_bar(stat = "identity") +
  labs(title = "Distinct Ids in Each Dataframe",
       x = "Dataframe",
       y = "Number of Distinct Ids") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the bar graph
print(bar_graph)

```

This graph gives us the clearest hint of what people are more likely to keep track of. The majority keep track of their daily activity and exercises. Slightly fewer keep track of their sleep, and even less than that monitor their heart rate. Very few measured their weight and recorded their fat etc.

```{r}
# Convert the date columns to a common format
daily_activity$ActivityDate <- as.Date(daily_activity$ActivityDate, format = "%m/%d/%Y")
sleep_day_cleaned$SleepDay <- as.Date(sleep_day_cleaned$SleepDay)

# Merge the two dataframes based on 'Id' and matching dates
merged_data <- inner_join(daily_activity, sleep_day_cleaned, by = c("Id", "ActivityDate" = "SleepDay"))

# Calculate the percentage of time spent in bed actually asleep
merged_data <- merged_data %>%
  mutate(PercentageTimeAsleep = (TotalMinutesAsleep / TotalTimeInBed) * 100)

# Select the relevant columns for the correlation matrix
selected_columns <- merged_data %>%
  select(
    VeryActiveDistance,
    Calories,
    SedentaryMinutes,
    TotalMinutesAsleep,
    PercentageTimeAsleep
  )

# Calculate the correlation matrix
correlation_matrix <- cor(selected_columns, use = "complete.obs")

corrplot(correlation_matrix, method = "color", type = "upper", tl.col = "black", tl.srt = 45)

# Print the correlation matrix
print(correlation_matrix)

```

From the correlation matrix above, we can see that burning a lot of calories seems to improve the "quality of sleep" and increasing sedentary minutes during the day decreases the total sleep time.

```{r}
# Convert date columns to a common format
daily_activity$ActivityDate <- as.Date(daily_activity$ActivityDate, format = "%m/%d/%Y")
sleep_day_cleaned$SleepDay <- as.Date(sleep_day_cleaned$SleepDay)
weight_log$Date <- as.POSIXct(weight_log$Date, format = "%Y-%m-%d %H:%M:%S")

# Join the dataframes based on 'Id' and matching dates
merged_data <- daily_activity %>%
  inner_join(sleep_day_cleaned, by = c("Id")) %>%
  inner_join(weight_log, by = c("Id"))

# Select the relevant columns
selected_columns <- merged_data %>%
  select(Calories, TotalMinutesAsleep, WeightKg, BMI)

# Create scatterplot with regression lines
scatterplot_calories_vs_minutes_asleep <- ggplot(selected_columns, aes(x = Calories, y = TotalMinutesAsleep)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add linear regression line
  labs(title = "Calories vs. Total Minutes Asleep",
       x = "Calories",
       y = "Total Minutes Asleep")

scatterplot_calories_vs_bmi <- ggplot(selected_columns, aes(x = Calories, y = BMI)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear regression line
  labs(title = "Calories vs. BMI",
       x = "Calories",
       y = "BMI")

# Display the scatterplots with regression lines
print(scatterplot_calories_vs_minutes_asleep)
print(scatterplot_calories_vs_bmi)

```

# End